---
title: "Annotations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Annotations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## QuPath-Compatible GeoJSON

wsiTools reads and writes GeoJSON annotations for pathology workflows.

```{r, purl = FALSE}
rois <- read_geojson("qupath_annotations.geojson")

viewer <- wsi_viewer_live(slide, tiled = TRUE)
viewer$open()
viewer$add_rois(rois)

edited <- viewer$get_rois()
write_geojson(edited, "edited_for_qupath.geojson")
```

The annotation system is designed for polygon and multipolygon ROIs, labels,
classes, colors, IDs, and basic measurements.

## Masks And Geometry

Mask images can be converted to ROIs, and ROIs can be rasterized back to masks
for CellPhenotyper, StarDist, Cellpose, Mesmer, or machine-learning pipelines.

```{r, purl = FALSE}
rois <- mask_to_rois("mask.png", class = "tumour")
mask <- rois_to_mask(rois, width = 2000, height = 1500)
```

## Measurements

```{r, purl = FALSE}
summarise_rois(rois)
measure_distance(c(10, 20), c(100, 200), pixel_size = 0.5)
measure_cell_density(cells, rois, pixel_size = 0.5)
```

More detail:

- [Examples gallery](https://github.com/tkcaccia/wsiTools/blob/main/docs/examples.md)
- [Project format](https://github.com/tkcaccia/wsiTools/blob/main/docs/projects.md)
