---
title: "Backends"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Backends}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## Runtime Capabilities

wsiTools is installable without heavy image libraries. Backends are optional
runtime capabilities discovered with:

```{r, purl = FALSE}
wsi_backends()
wsi_has_vips()
wsi_has_openslide()
wsi_has_native_czi()
wsi_has_bioformats()
```

## Backend Roles

| Backend | Main role |
| --- | --- |
| libvips | conversion, pyramids, thumbnails, fast tiles |
| OpenSlide | SVS/NDPI/SCN/MRXS-style WSI reading |
| native CZI | fast CZI scene/section region reads |
| Bio-Formats | broad microscopy fallback |
| ImageMagick | ordinary image fallback previews |
| httpuv | live R/browser bridge |

When a backend is unavailable, backend-dependent functions should fail with an
informative error instead of crashing R.

## Setup Reports

```{r, purl = FALSE}
wsi_setup_report()
wsi_diagnose()
```

More detail:

- [Backend setup guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/backends.md)
- [Installation guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/installation.md)
