---
title: "CellPhenotyper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CellPhenotyper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## Open A CellPhenotyper Project

CellPhenotyper projects can be opened from their output directory. wsiTools
uses `00_execution/project_outputs.tsv` to discover available outputs where
possible.

```{r, purl = FALSE}
project <- read_cellphenotyper_project(
  "/path/to/CellPhenotyper/results"
)

viewer <- viewer_cellphenotyper(project, live = TRUE)
viewer$open()
```

## Typical Layers

Depending on available outputs, the viewer may expose:

- input H&E or OME-TIFF images;
- GigaTIME/mIHC channel overlays;
- shifted or registered channel sources;
- cells, masks, centroids, or polygons;
- CellPhenotyper and GrandQC GeoJSON annotations;
- KODAMA/CellPhenotyper selection plots.

## Returned R Objects

```{r, purl = FALSE}
viewer$get_rois()
viewer$get_segmentation()
viewer$get_cell_summary()
viewer$get_channel_settings()
```

More detail:

- [Spatial omics guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/spatial-omics.md)
- [Examples gallery](https://github.com/tkcaccia/wsiTools/blob/main/docs/examples.md)
