---
title: "Concepts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Concepts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

wsiTools is built around a few concepts that explain most viewer behavior.
Understanding them makes it easier to diagnose slow images, missing overlays,
and coordinate alignment problems.

## Whole-Slide Images Are Not Ordinary Images

A whole-slide image can contain billions of pixels. wsiTools therefore avoids
reading the complete level-0 image into R memory. The normal workflow is:

1. read metadata;
2. choose a pyramid level or region;
3. request only the image tiles needed for the current viewport;
4. keep annotations, spots, cells, and measurements in slide coordinates.

## Static Viewer And Live Viewer

Use `wsi_viewer()` when you want an HTML file that can be opened later without
R. This is useful for sharing a read-only view.

Use `wsi_viewer_live()` when browser edits should return to R. The live viewer
starts a local `httpuv` service and opens a URL such as
`http://127.0.0.1:<port>/...`.

```{r, purl = FALSE}
slide <- wsi_open("/path/to/image.svs")

static_html <- wsi_viewer(slide)

viewer <- wsi_viewer_live(slide, tiled = TRUE)
viewer$open()
viewer$get_rois()
```

If the browser opens a `file://` page, R synchronization is not active.

## Tiles And OpenSeadragon

OpenSeadragon is the browser tile viewer. It does not decode raw SVS, CZI,
BTF, or OME-TIFF files by itself. Instead, it requests JPEG or PNG tiles.
wsiTools can provide those tiles in two ways:

- prebuilt Deep Zoom tiles for the smoothest static viewing;
- dynamic live tile URLs backed by libvips, OpenSlide, native CZI, or another
  available region reader.

This is why full-resolution viewing depends on a tile-capable backend.

## Coordinate Systems

All exported annotations and spatial overlays should use the original slide
coordinate system, not the current browser zoom level. Zooming, panning,
rotating, flipping, or opening a multi-view pane must not change the saved
GeoJSON coordinates.

Spatial omics data add a second layer of coordinates. For Seurat, Giotto, and
SpatialExperiment objects, wsiTools tries to keep the relation between
expression values and image coordinates explicit, especially when an object
contains multiple tissue sections.

## Image Formats And Backends

Format support is backend-dependent. The package can install without heavy
image tools, but advanced workflows need runtime capabilities.

| Format | Recommended backend | Status |
| --- | --- | --- |
| SVS | OpenSlide or libvips | supported when backend available |
| NDPI, SCN, MRXS | OpenSlide | backend-dependent |
| OME-TIFF, BTF, TIFF | libvips or Bio-Formats | backend-dependent |
| CZI | native CZI or Bio-Formats | improving |
| ordinary PNG/JPEG | ImageMagick or base image readers | fallback/previews |

Check the current machine with:

```{r, purl = FALSE}
wsi_backends()
wsi_diagnose(live_test = FALSE)
```

## Project State

A `.wsiproject` records the analysis state: slide paths, annotations,
trajectories, segmentation, measurements, channel settings, tile manifests,
viewport position, and provenance. It does not copy whole-slide image pixels by
default.

## Architecture

```{r, echo = FALSE, results = "asis"}
cat("```mermaid\n")
cat("flowchart TD\n")
cat("  R[\"R session\"] --> Bridge[\"httpuv / WebSocket / polling\"]\n")
cat("  Bridge --> Browser[\"Browser viewer\"]\n")
cat("  Browser --> OSD[\"OpenSeadragon tiles\"]\n")
cat("  OSD --> Backends[\"OpenSlide / libvips / native CZI / Bio-Formats\"]\n")
cat("  Browser --> Overlays[\"Annotations / spots / cells / trajectories\"]\n")
cat("  Overlays --> R\n")
cat("```\n")
```

For the detailed architecture guide, see
[wsiTools architecture](https://github.com/tkcaccia/wsiTools/blob/main/docs/architecture.md).

