---
title: "Desktop app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Desktop app}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

The optional wsiTools Desktop app is a Tauri launcher for users who do not want
to type R commands. It does not replace the R package. It finds `Rscript`,
starts R in the background, and opens the live wsiTools viewer.

## Install

1. Install R from CRAN.
2. Install wsiTools in R.
3. Download the desktop installer from GitHub Releases.
4. Start the app and choose **Open project** or **Create new project**.

```{r, purl = FALSE}
install.packages("remotes", repos = "https://cloud.r-project.org")
remotes::install_github("tkcaccia/wsiTools", upgrade = "never")

library(wsiTools)
wsi_diagnose(live_test = FALSE)
```

Downloads and platform notes:

- [Desktop downloads](https://github.com/tkcaccia/wsiTools/blob/main/docs/downloads.md)
- [Compile the Tauri app](https://github.com/tkcaccia/wsiTools/blob/main/docs/tauri-build.md)

## Create A New Project

The guided project workflow is:

1. add one or more microscopy images;
2. associate optional cell annotations, tissue annotations, and spatial omics
   objects to each image;
3. map tissue or sample names when a spatial object contains multiple sections;
4. click **Run R** to start the live viewer.

The viewer opens in a separate window. R remains responsible for image access,
dynamic tiles, synchronization, saving spatial objects, and backend checks.

## Common Problems

If the viewer opens white, open the app log and check whether R returned a
viewer URL. Typical causes are:

- `Rscript` is not on `PATH`;
- wsiTools is not installed in the R used by the app;
- libvips/OpenSlide/native CZI is missing for a full-resolution workflow;
- a selected spatial object does not contain coordinates for the selected
  tissue/sample.

Use:

```{r, purl = FALSE}
wsi_backends()
wsi_diagnose()
```

