---
title: "Getting started with wsiTools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with wsiTools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## What is a WSI?

A whole-slide image (WSI) is a digitized microscopy slide. In digital pathology
and bioimage analysis, a single file may contain billions of pixels, metadata,
multiple pyramid levels, and associated images such as labels or macro views.

## Why region-based reading is necessary

Loading an entire level-0 WSI into R memory is usually impractical. wsiTools is
therefore designed around:

- metadata reads
- pyramid levels
- explicit rectangular region reads
- tile-coordinate manifests
- backend-driven thumbnails and conversion
- streaming export where libvips supports it

## Opening slides

```{r, eval=FALSE, purl=FALSE}
library(wsiTools)

wsi_backends()

slide <- wsi_open("sample.svs")
slide
```

`wsi_open()` returns a lightweight `wsi_slide` object. It validates the file,
chooses a backend, and reads metadata. It does not load the full slide.

## Reading metadata

```{r, eval=FALSE, purl=FALSE}
wsi_info(slide)
wsi_levels(slide)
wsi_mpp(slide)
wsi_objective_power(slide)
```

The exact metadata depends on the file and backend. OpenSlide properties are
kept in `wsi_properties(slide)` for reproducible preprocessing pipelines.

## Reading patches

```{r, eval=FALSE, purl=FALSE}
patch <- wsi_read_region(
  slide,
  x = 10000,
  y = 20000,
  width = 512,
  height = 512,
  level = 0,
  format = "array"
)
```

Coordinates are level-0 coordinates. Width and height are output pixels at the
requested pyramid level, following the OpenSlide convention.

## Creating thumbnails

```{r, eval=FALSE, purl=FALSE}
thumb <- wsi_thumbnail(slide, width = 1000, format = "magick")
```

When libvips is available, thumbnails are generated through libvips rather than
by loading the full resolution image into R.

## Viewing slides interactively

```{r, eval=FALSE, purl=FALSE}
viewer <- wsi_viewer(slide, width = 1600)
```

`wsi_viewer()` writes a self-contained HTML preview with pan, zoom, and level-0
coordinate readout. It visualizes a backend-generated thumbnail, so it remains
safe for very large WSI files.

The viewer toolbar is organized into menus: `Navigate`, `Annotations`,
`GeoJSON`, `View`, and, when enabled, `Stains`. These menus group pan/select
modes, fit and 1:1 zoom, ROI and label toggles, ROI opacity, previous/next ROI
navigation, a side window listing all GeoJSON geometries, crosshair display,
coordinate copying, polygon drawing, and GeoJSON export. Use `GeoJSON` to open
the geometry list; each row shows the geometry type, bounds, point count,
source, and id. Use `Draw ROI`, click polygon vertices, double-click or press
`Finish`, then use `Save GeoJSON`. In `Brush` mode, painting with no selected
annotation creates a new ROI; painting with an annotation selected extends it;
holding `Alt` while brushing removes from the selected annotation. Use `Edit`
to move vertices, double-click an edge to insert a vertex, and Backspace/Delete
to remove the active vertex. The `Name`, `Class`, and `Custom class` controls
update the selected annotation label and category before GeoJSON export.
`Ctrl+Z` undoes annotation edits in the viewer, with the last 10 committed
annotation states retained.

For full-resolution zooming, use tiled mode. This asks libvips to create a
Deep Zoom tile pyramid on disk, then writes an HTML viewer that requests the
appropriate tile resolution as you zoom.

```{r, eval=FALSE, purl=FALSE}
viewer <- wsi_viewer(
  slide,
  mode = "tiles",
  output = "sample_viewer.html",
  tile_dir = "sample_viewer_tiles",
  tile_size = 512
)
```

To compare H&E/IHC, original/processed images, serial sections, or WSI/mask
views, use the side-by-side viewer. Zoom, pan, and cursor position are linked
when `sync = TRUE`.

```{r, eval=FALSE, purl=FALSE}
viewer_compare(
  "sample_he.svs",
  "sample_ihc.svs",
  sync = TRUE,
  output = "he_ihc_compare.html"
)
```

OME-Zarr directories can be opened as metadata-backed slide handles. This
extracts dimensions and pyramid levels without decoding all chunks.

```{r, eval=FALSE, purl=FALSE}
zarr <- open_omezarr("sample.ome.zarr")
omezarr_metadata("sample.ome.zarr")
```

## Deconvolving IHC stains

For hematoxylin plus HRP/DAB immunohistochemistry, `wsiTools` can deconvolve a
requested patch into two stain channels:

```{r, eval=FALSE, purl=FALSE}
channels <- wsi_deconvolve_region(
  slide,
  x = 10000,
  y = 20000,
  width = 1024,
  height = 1024
)

str(channels$hematoxylin)
str(channels$hrp)
```

The interactive IHC viewer performs the same H-DAB deconvolution on the visible
canvas pixels. With `mode = "tiles"`, zooming still requests Deep Zoom tiles
rather than loading the full slide into R:

```{r, eval=FALSE, purl=FALSE}
ihc_viewer <- wsi_viewer_ihc(
  slide,
  mode = "tiles",
  output = "sample_ihc_viewer.html",
  tile_dir = "sample_viewer_tiles"
)
```

The `Stains` menu provides an `IHC` toggle, separate hematoxylin and HRP/DAB
channel visibility controls, color pickers, and gain sliders.

For brightfield multiplex IHC, create selectable stain channels and pass them
to the multi-channel viewer:

```{r, eval=FALSE, purl=FALSE}
multi_channels <- wsi_stain_channels(
  name = c("Hematoxylin", "HRP/DAB", "Fast Red"),
  vector = list(
    c(0.650, 0.704, 0.286),
    c(0.268, 0.570, 0.776),
    c(0.213, 0.851, 0.477)
  ),
  colour = c("#4b3f99", "#8b5a2b", "#d73027")
)

multi_viewer <- wsi_viewer_multi_ihc(
  slide,
  channels = multi_channels,
  mode = "tiles",
  output = "sample_multi_ihc_viewer.html",
  tile_dir = "sample_viewer_tiles"
)
```

The `Stains` menu adds the `mIHC` toggle plus a checkbox, colour picker, and
gain slider for every channel. RGB brightfield deconvolution can separate up to
three independent stain vectors at once, so use assay-specific vectors for the
markers you want to inspect.

## Generating tile grids

```{r, eval=FALSE, purl=FALSE}
grid <- wsi_tile_grid(
  slide,
  tile_size = 512,
  overlap = 64,
  level = 0
)

head(grid)
```

`wsi_tile_grid()` reads no image pixels. It only returns tile coordinates,
dimensions, downsample factors, and suggested filenames.

## Extracting tiles

```{r, eval=FALSE, purl=FALSE}
manifest <- wsi_tile(
  slide,
  output_dir = "tiles",
  tile_size = 512,
  overlap = 0,
  level = 0,
  format = "png"
)
```

The returned manifest records file paths and coordinates for downstream machine
learning or bioinformatics workflows.

If you already have tile locations from a CSV file, a model, or an external
annotation workflow, use them directly:

```{r, eval=FALSE, purl=FALSE}
coords <- data.frame(
  tile_id = c("spot_001", "spot_002"),
  x = c(10000, 12500),
  y = c(20000, 21500)
)

coord_grid <- wsi_tile_grid_from_coords(
  slide,
  coords,
  tile_size = 512,
  anchor = "center"
)

manifest <- wsi_tile_from_coords(
  slide,
  coords,
  output_dir = "coordinate_tiles",
  tile_size = 512,
  anchor = "center",
  bounds = "trim"
)
```

Coordinates are level-0 slide pixels. Set `anchor = "top_left"` when `x` and
`y` are the tile corner, or `anchor = "center"` when they mark the center.

For stride-based extraction, `extract_tiles()` returns coordinates unless an
output directory is supplied.

```{r, eval=FALSE, purl=FALSE}
grid <- extract_tiles(slide, tile_size = 512, stride = 256, save_images = FALSE)
manifest <- extract_tiles(slide, tile_size = 512, stride = 512, output_dir = "tiles")
```

## Cropping ROIs

```{r, eval=FALSE, purl=FALSE}
roi <- wsi_read_geojson("annotations.geojson")
wsi_crop_roi(slide, roi, output_dir = "roi_crops", format = "tiff")
```

The first milestone supports QuPath-style GeoJSON parsing and ROI bounding
boxes. Polygon-aware ROI operations require the optional `sf` package and will
be expanded in future releases.

GeoJSON annotations can also be drawn over the interactive viewer:

```{r, eval=FALSE, purl=FALSE}
roi <- wsi_read_geojson("annotations.geojson")

viewer <- wsi_viewer_roi(
  slide,
  roi,
  mode = "tiles",
  output = "sample_roi_viewer.html",
  tile_dir = "sample_viewer_tiles"
)
```

ROI overlay coordinates are interpreted as level-0 slide pixels, matching
QuPath-style GeoJSON exports.

ROIs can be relabelled and written back to GeoJSON for use in QuPath or other
pathology pipelines.

```{r, eval=FALSE, purl=FALSE}
roi <- read_geojson("annotations.geojson")
roi <- wsi_set_roi_class(roi, "tumour", roi_id = roi$roi_id[1])
write_geojson(roi, "annotations_relabelled.geojson", overwrite = TRUE)
```

Optional external segmentation tools remain outside wsiTools dependencies. A
typical workflow is to export an ROI crop, run an external model, import its
output, and display the result as an overlay.

```{r, eval=FALSE, purl=FALSE}
export_roi_crop(slide, roi, "roi_crop.png", roi_id = roi$roi_id[1])
segmentation <- import_segmentation("model_output.geojson")
viewer_add_segmentation(slide, segmentation, output = "segmentation_overlay.html")
```

Basic measurements and class summaries use ROI coordinates and optional pixel
size metadata.

```{r, eval=FALSE, purl=FALSE}
cells <- data.frame(x = c(1000, 1200, 3000), y = c(800, 900, 1800))
measure_cell_density(cells, roi, pixel_size = wsi_mpp(slide))
summarise_rois(roi, cells = cells, pixel_size = wsi_mpp(slide), file = "roi_summary.csv")
```

For serial sections, estimate a simple affine transform from manual landmarks
and transfer ROIs.

```{r, eval=FALSE, purl=FALSE}
transform <- estimate_transform(landmarks1, landmarks2)
transformed_roi <- transform_rois(roi, transform)
```

## Converting to pyramidal TIFF/OME-TIFF

```{r, eval=FALSE, purl=FALSE}
wsi_convert(
  input = "sample.svs",
  output = "sample.ome.tiff",
  format = "ome-tiff",
  pyramid = TRUE,
  compression = "lzw"
)
```

Conversion is handled by libvips when available. wsiTools constructs command
arguments safely and never overwrites existing files unless
`overwrite = TRUE`.

## Practical notes on OpenSlide and libvips

OpenSlide is the preferred first backend for many pathology WSI formats such as
Aperio SVS, Hamamatsu NDPI, Leica SCN, MIRAX MRXS, and related pyramidal files,
subject to local backend support.

libvips is used for efficient thumbnails, cropping, conversion, and pyramidal
TIFF export. OME-TIFF behavior depends on the installed libvips version and its
TIFF writer capabilities.

Always check:

```{r, eval=FALSE, purl=FALSE}
wsi_backends()
```

Close slide handles when finished:

```{r, eval=FALSE, purl=FALSE}
wsi_close(slide)
```
