---
title: "Getting started with wsiTools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with wsiTools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## What is a WSI?

A whole-slide image (WSI) is a digitized microscopy slide. In digital pathology
and bioimage analysis, a single file may contain billions of pixels, metadata,
multiple pyramid levels, and associated images such as labels or macro views.

## Why region-based reading is necessary

Loading an entire level-0 WSI into R memory is usually impractical. wsiTools is
therefore designed around:

- metadata reads
- pyramid levels
- explicit rectangular region reads
- tile-coordinate manifests
- backend-driven thumbnails and conversion
- streaming export where libvips supports it

## Installation and setup

Install the lightweight package first, then check which optional WSI backends
are available on your machine.

```{r, eval=FALSE}
install.packages("remotes", repos = "https://cloud.r-project.org")
remotes::install_github(
  "tkcaccia/wsiTools",
  upgrade = "never",
  build_vignettes = FALSE,
  force = TRUE
)

library(wsiTools)
packageVersion("wsiTools")
wsi_backends()
```

The current GitHub package version is `0.1.9`. If R reports an older version
after installation, remove the old package directory and reinstall from
`tkcaccia/wsiTools` without pinning an old commit.

OpenSlide, libvips, StarDist, Cellpose, Bio-Formats, and live web services are
runtime capabilities rather than mandatory installation dependencies. Use
`wsi_setup()` after installation to print a setup report with copyable commands
for optional tools.

```{r, eval=FALSE}
# Check what is available and what is missing
wsi_setup()

# Preview the backend install plan without running system commands
wsi_install_backends(install = FALSE)

# After reviewing the plan, optionally run supported installers.
# In interactive R, wsiTools asks before running commands.
wsi_install_backends(method = "auto")
```

Examples for common platforms:

```{r, eval=FALSE}
# macOS with Homebrew
wsi_install_backends(method = "homebrew")

# Ubuntu/Debian. sudo is never used unless you opt in.
wsi_install_backends(method = "apt", allow_sudo = TRUE)

# Fedora
wsi_install_backends(method = "dnf", allow_sudo = TRUE)

# Windows. OpenSlide is usually installed separately through conda-forge,
# MSYS2/vcpkg, or official binaries and then added to PATH.
wsi_install_backends(
  tools = c("libvips", "imagemagick"),
  method = "winget"
)

# Conda or mamba environment
wsi_install_backends(
  tools = c("openslide", "libvips", "imagemagick", "bioformats"),
  method = "conda"
)
```

Restart R or RStudio after installing system tools, then run `wsi_backends()`
again. StarDist and Cellpose are external segmentation tools; keep them in
separate environments and enable them explicitly rather than making them
mandatory package dependencies.

### Bio-Formats Java helper and command-line tools

For first visualization of CZI files, wsiTools should use a tile/region reader,
not `bfconvert`. wsiTools first tries OpenSlide and libvips for CZI variants
they can already read, because those backends are fast and tile-friendly. The
next preferred path is the native CZI bridge. A second optional path is the
Bio-Formats Java helper, which uses a tiny Java class around Bio-Formats
`ImageReader` region reads.

```{r, eval=FALSE}
Sys.setenv(WSITOOLS_BIOFORMATS_JAR = "/path/to/bioformats_package.jar")
# Optional if java/javac are not on PATH:
# Sys.setenv(WSITOOLS_JAVA = "/path/to/java")
# Sys.setenv(WSITOOLS_JAVAC = "/path/to/javac")

wsi_backends()
wsi_viewer_project("sample.czi")
```

The helper compiles into R's temporary directory and reads only requested
regions or low-resolution scenes. It does not convert the whole file for first
opening.

Bio-Formats command-line support means the command-line tools are visible to R.
The commands wsiTools checks for are `showinf` and `bfconvert`; these remain
useful for metadata and conversion workflows.

The recommended conda route is:

```{bash, eval=FALSE}
conda create -n wsitools-bioformats --override-channels -c ome -c conda-forge bftools
conda activate wsitools-bioformats
showinf -version
bfconvert -version
```

`--override-channels` tells conda to use only the explicitly listed channels.
This avoids Anaconda default-channel Terms of Service prompts on managed
Windows computers.

If R is running inside the active conda environment, this can also be requested
from R:

```{r, eval=FALSE}
wsi_install_backends(tools = "bioformats", method = "conda")
wsi_backends()
```

Manual installation is also possible: download `bftools.zip` from the official
OME Bio-Formats downloads page, unzip it into a permanent folder, add that
folder to `PATH`, then restart R/RStudio. Installing the Bio-Formats plugin in
Fiji/ImageJ alone does not usually expose `showinf` and `bfconvert` to R.

On Windows, if a previous installation was interrupted, remove stale lock
directories before retrying:

```{r, eval=FALSE}
lib <- Sys.getenv("R_LIBS_USER")
unlink(file.path(lib, "00LOCK-wsiTools"), recursive = TRUE, force = TRUE)
unlink(file.path(lib, "wsiTools"), recursive = TRUE, force = TRUE)

remotes::install_github(
  "tkcaccia/wsiTools",
  upgrade = "never",
  build_vignettes = FALSE,
  force = TRUE,
  INSTALL_opts = "--no-multiarch"
)
```

## Opening slides

```{r, eval=FALSE, purl=FALSE}
library(wsiTools)

wsi_backends()

slide <- wsi_open("sample.svs")
slide
```

`wsi_open()` returns a lightweight `wsi_slide` object. It validates the file,
chooses a backend, and reads metadata. It does not load the full slide.
OpenSlide, libvips, StarDist, Cellpose, `httpuv`, and `callr` are optional
runtime capabilities; the package should still install and load without them.

For Zeiss CZI files on Windows, ImageMagick is not a CZI backend. wsiTools first
tries OpenSlide and libvips; if neither can read the CZI variant, use the native
CZI tile/region reader for visualization by installing ZEISS libCZI/libCZIAPI
and setting `WSITOOLS_LIBCZIAPI` when needed. Alternatively, configure
`WSITOOLS_BIOFORMATS_JAR` for the Bio-Formats Java helper. Bio-Formats
command-line tools remain useful for metadata and conversion, but wsiTools does
not run `bfconvert` automatically for first visualization.

For first visualization with the native bridge:

```{r, eval=FALSE, purl=FALSE}
# Inspect the commands first; nothing is installed in this call.
wsi_install_native_czi(install = FALSE)

# Build ZEISS libCZI/libCZIAPI into the user cache after reviewing
# and accepting the libCZI license notice. Requires git, cmake, and
# a working C/C++ toolchain.
wsi_install_native_czi(
  accept_license = TRUE,
  ask = FALSE,
  persist = TRUE
)

wsi_has_native_czi()

wsi_viewer_project("sample.czi")
```

If you already installed libCZIAPI manually, point wsiTools at the shared
library instead:

```{r, eval=FALSE, purl=FALSE}
Sys.setenv(WSITOOLS_LIBCZIAPI = "C:/path/to/libCZIAPI.dll")
```

The project viewer opens CZI files with a low-resolution scene preview first.
When libCZI reports pyramid layers, wsiTools chooses a downsampled layer for the
first overview instead of immediately generating a high-resolution browser
preview. You can tune the first overview size with
`WSITOOLS_CZI_INITIAL_PREVIEW_WIDTH` and `WSITOOLS_CZI_MIN_PREVIEW_WIDTH`.

The old Python wrapper route is kept only as a legacy fallback. To use it
explicitly, set `WSITOOLS_CZI_ALLOW_PYTHON=true` and configure
`WSITOOLS_CZI_PYTHON`.

For metadata only, install Bio-Formats command-line tools and open explicitly
with `backend = "bioformats"`.

## Reading metadata

```{r, eval=FALSE, purl=FALSE}
wsi_info(slide)
wsi_levels(slide)
wsi_mpp(slide)
wsi_objective_power(slide)
```

The exact metadata depends on the file and backend. OpenSlide properties are
kept in `wsi_properties(slide)` for reproducible preprocessing pipelines.

## Reading patches

```{r, eval=FALSE, purl=FALSE}
patch <- wsi_read_region(
  slide,
  x = 10000,
  y = 20000,
  width = 512,
  height = 512,
  level = 0,
  format = "array"
)
```

Coordinates are level-0 coordinates. Width and height are output pixels at the
requested pyramid level, following the OpenSlide convention.

## Creating thumbnails

```{r, eval=FALSE, purl=FALSE}
thumb <- wsi_thumbnail(slide, width = 1000, format = "magick")
```

When libvips is available, thumbnails are generated through libvips rather than
by loading the full resolution image into R.

## Viewing slides interactively

```{r, eval=FALSE, purl=FALSE}
viewer <- wsi_viewer(slide, width = 1600)
```

`wsi_viewer()` writes a self-contained HTML preview with pan, zoom, and level-0
coordinate readout. It visualizes a backend-generated thumbnail, so it remains
safe for very large WSI files.

The viewer toolbar is organized into menus: `Navigate`, `Annotations`,
`GeoJSON`, `View`, and, when enabled, `Stains`. These menus group pan/select
modes, fit and 1:1 zoom, ROI and label toggles, ROI opacity, previous/next ROI
navigation, a side window listing all GeoJSON geometries, crosshair display,
coordinate copying, polygon drawing, and GeoJSON export. Use `GeoJSON` to open
the geometry list; each row shows the geometry type, bounds, point count,
source, and id. Use `Draw ROI`, click polygon vertices, double-click or press
`Finish`, then use `Save GeoJSON`. In `Brush` mode, each normal stroke creates
a new automatically named annotation. If the painted area touches an existing
annotation with the same label, the regions merge; holding `Alt` on
Windows/Linux or `Command` on macOS while brushing removes from the selected
annotation. The compact ROI report does not open
automatically; select an ROI and click `ROI summary` when you want to inspect
area, cell count, and density. Use `Edit`
to move vertices, double-click an edge to insert a vertex, and Backspace/Delete
to remove the active vertex. Drawn, painted, and edited annotation regions are
kept class-exclusive: annotations with the same class label merge into one
multi-part annotation, while areas already occupied by a different class label
are clipped from the new or refined annotation. The Class Presets and History sections in the
annotation panel can be maximized with their `Maximize` buttons and restored
with `Esc` or `Restore`. The toolbar `Class`, `Custom class`, and `Set next
class` controls set the class for the next drawn or painted annotation;
selecting an ROI does not overwrite these controls, and changing them does not
relabel the selected annotation. Use the annotation manager controls to change
an existing annotation before GeoJSON export.
Annotation colors are category-driven, so imported, drawn, and painted ROIs
with the same class label share the same class color. `Ctrl+Z` undoes
annotation edits in the viewer, with the last 10 committed
annotation states retained.

For full-resolution zooming, use tiled mode. OpenSeadragon is a browser tile
viewer: it needs browser-readable tiles, not raw SVS, BTF, or OME-TIFF pixels.
The static viewer asks libvips to create a Deep Zoom tile pyramid on disk, then
writes an HTML viewer that requests the appropriate tile resolution as you
zoom. This static viewer has no automatic feedback channel to R after it is
opened.

```{r, eval=FALSE, purl=FALSE}
viewer <- wsi_viewer(
  slide,
  mode = "tiles",
  output = "sample_viewer.html",
  tile_dir = "sample_viewer_tiles",
  tile_size = 512
)
```

For R-interconnected work, use the live viewer. It starts a local `httpuv`
bridge; WebSocket is used when available and HTTP POST/polling remains the
fallback. Live mode can either use the same precomputed Deep Zoom tiles or a
dynamic tile server that generates and caches tiles on demand from region
reads:

```{r, eval=FALSE, purl=FALSE}
viewer <- wsi_viewer_live(
  slide,
  mode = "tiles",
  dynamic_tiles = TRUE,
  dynamic_tile_format = "jpg",
  transport = "auto",
  wait = FALSE
)

viewer$get_rois()
viewer$get_measurements()
viewer$get_channel_settings()
```

The dynamic endpoint uses URLs like
`/tiles/{slide_id}/{level}/{x}/{y}.jpg`. It reads only requested regions
through libvips/OpenSlide where available and removes its temporary tile cache
when the live session is stopped.

The same live tiled architecture can overlay mIHC or probability-map OME-TIFF
pages on a registered H&E slide. The H&E slide remains the base image; each
mIHC page becomes an independent OpenSeadragon channel layer with visibility,
opacity, colour, gain, and contrast controls.

```{r, eval=FALSE, purl=FALSE}
viewer <- wsi_viewer_he_mihc(
  he = "sample_he.svs",
  mihc = "sample_mihc_probs.ome.tif",
  registration = "shift.json",
  dynamic_tiles = FALSE,
  tile_dir = "sample_he_deepzoom",
  transport = "auto",
  wait = FALSE
)

viewer$get_channel_settings()
```

For H&E plus mIHC overlays, precomputed H&E Deep Zoom tiles are recommended for
fluid pan/zoom. Use `dynamic_tiles = TRUE` only when a tile pyramid is not
available; in that fallback mode `wsi_viewer_he_mihc()` uses JPEG base tiles by
default.

To compare H&E/IHC, original/processed images, serial sections, or WSI/mask
views, use the side-by-side viewer. Zoom, pan, and cursor position are linked
when `sync = TRUE`.

```{r, eval=FALSE, purl=FALSE}
viewer_compare(
  "sample_he.svs",
  "sample_ihc.svs",
  sync = TRUE,
  output = "he_ihc_compare.html"
)
```

OME-Zarr directories can be opened as metadata-backed slide handles. This
extracts dimensions, axes, chunks, dtype, coordinate transforms, and pyramid
levels without decoding all chunks.

```{r, eval=FALSE, purl=FALSE}
zarr <- open_omezarr("sample.ome.zarr")
meta <- omezarr_metadata("sample.ome.zarr")
meta$axis_table
meta$levels
```

## Deconvolving IHC stains

For hematoxylin plus HRP/DAB immunohistochemistry, `wsiTools` can deconvolve a
requested patch into two stain channels:

```{r, eval=FALSE, purl=FALSE}
channels <- wsi_deconvolve_region(
  slide,
  x = 10000,
  y = 20000,
  width = 1024,
  height = 1024
)

str(channels$hematoxylin)
str(channels$hrp)
```

The interactive IHC viewer performs the same H-DAB deconvolution on the visible
canvas pixels. With `mode = "tiles"`, zooming still requests Deep Zoom tiles
rather than loading the full slide into R:

```{r, eval=FALSE, purl=FALSE}
ihc_viewer <- wsi_viewer_ihc(
  slide,
  mode = "tiles",
  output = "sample_ihc_viewer.html",
  tile_dir = "sample_viewer_tiles"
)
```

The `Stains` menu provides an `IHC` toggle, separate hematoxylin and HRP/DAB
channel visibility controls, color pickers, opacity, gain, and contrast-window
sliders. These display settings are synced back to R in live sessions through
`viewer$get_channel_settings()`.

For H&E images, use the H&E stain helpers on a patch, tile, or thumbnail. The
functions separate hematoxylin, eosin, and residual staining in
optical-density space and avoid full-slide reads:

```{r, eval=FALSE, purl=FALSE}
patch <- wsi_read_region(
  slide,
  x = 10000,
  y = 20000,
  width = 1024,
  height = 1024
)

he_channels <- wsi_deconvolve_he(patch)
names(he_channels)
#> hematoxylin, eosin, residual

stain_matrix <- wsi_estimate_stain_matrix(patch, method = "macenko")

normalized <- wsi_normalize_stains(
  patch,
  method = "macenko",
  target_matrix = wsi_he_stain_matrix()
)
```

The same three-channel H&E separation is available in the interactive viewer:

```{r, eval=FALSE, purl=FALSE}
he_viewer <- wsi_viewer_he(
  slide,
  mode = "tiles",
  output = "he_stain_viewer.html"
)
```

There is also a region wrapper for WSI preprocessing pipelines:

```{r, eval=FALSE, purl=FALSE}
he_region <- wsi_deconvolve_he_region(
  slide,
  x = 10000,
  y = 20000,
  width = 1024,
  height = 1024
)

normalized_region <- wsi_stain_normalize_region(
  slide,
  x = 10000,
  y = 20000,
  width = 1024,
  height = 1024,
  method = "macenko"
)
```

`method = "vahadane"` provides a small dependency-free NMF approximation for
Vahadane-style separation. For production studies, validate stain vectors and
normalisation settings on reference material from the same laboratory workflow.

For brightfield multiplex IHC, create selectable stain channels and pass them
to the multi-channel viewer:

```{r, eval=FALSE, purl=FALSE}
multi_channels <- wsi_stain_channels(
  name = c("Hematoxylin", "HRP/DAB", "Fast Red"),
  vector = list(
    c(0.650, 0.704, 0.286),
    c(0.268, 0.570, 0.776),
    c(0.213, 0.851, 0.477)
  ),
  colour = c("#4b3f99", "#8b5a2b", "#d73027")
)

multi_viewer <- wsi_viewer_multi_ihc(
  slide,
  channels = multi_channels,
  mode = "tiles",
  output = "sample_multi_ihc_viewer.html",
  tile_dir = "sample_viewer_tiles"
)
```

The `Stains` menu adds the `mIHC` toggle plus a checkbox, colour picker, and
gain slider for every channel. RGB brightfield deconvolution can separate up to
three independent stain vectors at once, so use assay-specific vectors for the
markers you want to inspect.

## Generating tile grids

```{r, eval=FALSE, purl=FALSE}
grid <- wsi_tile_grid(
  slide,
  tile_size = 512,
  overlap = 64,
  level = 0
)

head(grid)
```

`wsi_tile_grid()` reads no image pixels. It only returns tile coordinates,
dimensions, downsample factors, and suggested filenames.

## Extracting tiles

```{r, eval=FALSE, purl=FALSE}
manifest <- wsi_tile(
  slide,
  output_dir = "tiles",
  tile_size = 512,
  overlap = 0,
  level = 0,
  format = "png"
)
```

The returned manifest records file paths and coordinates for downstream machine
learning or bioinformatics workflows.

If you already have tile locations from a CSV file, a model, or an external
annotation workflow, use them directly:

```{r, eval=FALSE, purl=FALSE}
coords <- data.frame(
  tile_id = c("spot_001", "spot_002"),
  x = c(10000, 12500),
  y = c(20000, 21500)
)

coord_grid <- wsi_tile_grid_from_coords(
  slide,
  coords,
  tile_size = 512,
  anchor = "center"
)

manifest <- wsi_tile_from_coords(
  slide,
  coords,
  output_dir = "coordinate_tiles",
  tile_size = 512,
  anchor = "center",
  bounds = "trim"
)
```

Coordinates are level-0 slide pixels. Set `anchor = "top_left"` when `x` and
`y` are the tile corner, or `anchor = "center"` when they mark the center.

For stride-based extraction, `extract_tiles()` returns coordinates unless an
output directory is supplied.

```{r, eval=FALSE, purl=FALSE}
grid <- extract_tiles(slide, tile_size = 512, stride = 256, save_images = FALSE)
manifest <- extract_tiles(slide, tile_size = 512, stride = 512, output_dir = "tiles")
```

For machine-learning patch extraction, `extract_tiles()` can work directly from
ROI annotations, balance classes, skip background, save a CSV manifest, and add
a reproducible train/validation split:

```{r, eval=FALSE, purl=FALSE}
manifest <- extract_tiles(
  slide,
  roi = roi,
  output_dir = "ml_tiles",
  tile_size = 512,
  stride = 512,
  sampling = "balanced",
  tiles_per_class = 500,
  skip_background = TRUE,
  split = "train_validation",
  train_fraction = 0.8,
  seed = 2026,
  manifest_file = "ml_tiles/manifest.csv"
)
```

Artifact filtering can be added as an optional quality-control pass. The
package reads candidate tiles one region at a time and appends flags for blur,
out-of-focus appearance, pen-like marks, fold-like dark saturated regions,
bubble-like bright regions, and very bright or dark tiles. Whitespace/background
labelling can be enabled at the same time and adds `whitespace_*` and
`background_*` manifest columns:

```{r, eval=FALSE, purl=FALSE}
manifest <- extract_tiles(
  slide,
  roi = roi,
  tile_size = 512,
  stride = 512,
  whitespace_filter = TRUE,
  whitespace_action = "flag",
  artifact_filter = TRUE,
  artifact_action = "flag",
  save_images = FALSE
)

clean_manifest <- subset(manifest, !artifact_flag & !whitespace_flag)
```

## Cropping ROIs

```{r, eval=FALSE, purl=FALSE}
roi <- wsi_read_geojson("annotations.geojson")
wsi_crop_roi(slide, roi, output_dir = "roi_crops", format = "tiff")
```

The first milestone supports QuPath-style GeoJSON parsing and ROI bounding
boxes. Polygon-aware ROI operations require the optional `sf` package and will
be expanded in future releases.

GeoJSON annotations can also be drawn over the interactive viewer:

```{r, eval=FALSE, purl=FALSE}
roi <- wsi_read_geojson("annotations.geojson")

viewer <- wsi_viewer_roi(
  slide,
  roi,
  mode = "tiles",
  output = "sample_roi_viewer.html",
  tile_dir = "sample_viewer_tiles"
)
```

ROI overlay coordinates are interpreted as level-0 slide pixels, matching
QuPath-style GeoJSON exports.

ROIs can be relabelled and written back to GeoJSON for use in QuPath or other
pathology pipelines. Polygon and multipolygon annotations retain QuPath object
ids, object type, classification labels and colors, measurements, lock state,
raw properties, and top-level GeoJSON metadata where possible.

```{r, eval=FALSE, purl=FALSE}
roi <- read_geojson("annotations.geojson")
roi <- wsi_set_roi_class(roi, "tumour", roi_id = roi$roi_id[1])
write_geojson(roi, "annotations_relabelled.geojson", overwrite = TRUE)
```

The same objects can round-trip through a live viewer session:

```{r, eval=FALSE, purl=FALSE}
viewer <- wsi_viewer_live(slide, wait = FALSE)
viewer$capabilities()

rois <- read_geojson("qupath_annotations.geojson")
viewer$add_rois(rois)

edited <- viewer$get_rois()
write_geojson(edited, "edited_for_qupath.geojson", overwrite = TRUE)
```

Annotation or segmentation masks can also be converted to ROI polygons. This
works best for already-small mask images or matrices; it should not be used to
load an entire WSI into memory.

```{r, eval=FALSE, purl=FALSE}
mask_rois <- read_mask_annotations(
  "tumour_mask.png",
  threshold = 0.5,
  class_map = c("1" = "tumour"),
  origin = c(x = 0, y = 0),
  scale = c(x = 1, y = 1)
)
viewer$add_rois(mask_rois)
write_geojson(mask_rois, "tumour_mask_annotations.geojson", overwrite = TRUE)
```

Optional external segmentation tools remain outside wsiTools dependencies. A
typical workflow is to export an ROI crop, run an external model, import its
output, and display the result as an overlay.

```{r, eval=FALSE, purl=FALSE}
export_roi_crop(slide, roi, "roi_crop.png", roi_id = roi$roi_id[1])
segmentation <- import_segmentation("model_output.geojson")
viewer_add_segmentation(slide, segmentation, output = "segmentation_overlay.html")
```

For live sessions, long-running operations can run in a background R process
when the optional `callr` package is installed. The viewer remains responsive,
and results are collected later through a small job object.

```{r, eval=FALSE, purl=FALSE}
# Preview tiles first as a locked viewer layer, then export the confirmed grid.
preview <- viewer$preview_tiles(tile_size = 512, stride = 512)
tiles <- viewer$extract_tile_preview("confirmed_tiles", format = "png")

job <- viewer$run_segmentation_async()
job$status()
cells <- job$result()

tile_job <- viewer$run_tiles_async(tile_size = 512, stride = 512, save_images = FALSE)
tiles <- tile_job$result()
```

Basic measurements and class summaries use ROI coordinates and optional pixel
size metadata. The report helper bundles ROI areas, class summaries,
cell-density tables, nearest-neighbour distances, boundary distances, and
optional hematoxylin/HRP-DAB intensity summaries into CSV-ready tables.

```{r, eval=FALSE, purl=FALSE}
cells <- data.frame(x = c(1000, 1200, 3000), y = c(800, 900, 1800))
measure_cell_density(cells, roi, pixel_size = wsi_mpp(slide))
summarise_rois(roi, cells = cells, pixel_size = wsi_mpp(slide), file = "roi_summary.csv")

patch_channels <- wsi_deconvolve_region(
  slide,
  x = 10000,
  y = 20000,
  width = 2048,
  height = 2048
)

ihc_roi <- measure_ihc_intensity(
  patch_channels,
  rois = roi,
  image_origin = c(x = 10000, y = 20000),
  dab_threshold = 0.1,
  pixel_size = wsi_mpp(slide)
)
ihc_class <- summarise_ihc_intensity(ihc_roi)

cell_counts <- wsi_cell_counts(
  slide,
  segmentation = cells,
  channels = patch_channels,
  rois = roi,
  image_origin = c(x = 10000, y = 20000),
  positive_threshold = c(hrp_dab = 0.1),
  pixel_size = wsi_mpp(slide),
  output_dir = "cell_counts",
  overwrite = TRUE
)

report <- measurement_report(
  roi,
  cells = cells,
  stains = patch_channels,
  image_origin = c(x = 10000, y = 20000),
  pixel_size = wsi_mpp(slide),
  output_dir = "measurement_report",
  prefix = "case_001"
)
```

Use `wsi_project()` to save a reopenable analysis session. The project stores a
small JSON index plus GeoJSON/CSV sidecar files for annotations, viewer state,
measurements, segmentation overlays, stain settings, and tile manifests.

```{r, eval=FALSE, purl=FALSE}
project <- wsi_project(
  "case_001.wsiproject",
  slide = slide,
  viewer_state = session,
  rois = roi,
  measurements = report,
  segmentation = segmentation,
  stains = patch_channels,
  tile_manifest = manifest,
  metadata = list(case_id = "case_001")
)

reopened <- wsi_read_project("case_001.wsiproject")
reopened$rois
reopened$tile_manifest

case_report <- wsi_case_report(
  reopened,
  output_dir = "case_001_report",
  overwrite = TRUE
)
case_report$html
case_report$files
```

For serial sections, estimate a simple affine transform from manual landmarks
and transfer ROIs.

```{r, eval=FALSE, purl=FALSE}
transform <- estimate_transform(landmarks1, landmarks2)
transformed_roi <- transform_rois(roi, transform)
```

## Converting to pyramidal TIFF/OME-TIFF

```{r, eval=FALSE, purl=FALSE}
wsi_convert(
  input = "sample.svs",
  output = "sample.ome.tiff",
  format = "ome-tiff",
  pyramid = TRUE,
  compression = "lzw",
  tile_size = 512,
  depth = "onepixel",
  region_shrink = "mean",
  predictor = "horizontal"
)

wsi_export_ome_tiff(
  input = "sample.svs",
  output = "sample.ome.tiff",
  compression = "lzw",
  overwrite = TRUE
)
```

Conversion is handled by libvips when available. wsiTools constructs command
arguments safely and never overwrites existing files unless
`overwrite = TRUE`. OME-TIFF export enables tiled pyramids, SubIFDs, BigTIFF,
and TIFF metadata properties by default; exact codec support depends on the
installed libvips/libtiff build.

## Practical notes on OpenSlide and libvips

The CRAN-facing core of wsiTools is intentionally lightweight. Installation
does not require OpenSlide, libvips, Python, StarDist, Cellpose, or a running web
service. Those are optional runtime capabilities that are checked when a user
asks for a backend-dependent operation.

OpenSlide is the preferred first backend for many pathology WSI formats such as
Aperio SVS, Hamamatsu NDPI, Leica SCN, MIRAX MRXS, and related pyramidal files,
subject to local backend support.

libvips is used for efficient thumbnails, cropping, conversion, and pyramidal
TIFF/OME-TIFF export. OME-TIFF behavior depends on the installed libvips
version and its TIFF writer capabilities.

StarDist and Cellpose integrations should remain external-command or local
service workflows: export an ROI crop, run the model in the user's chosen
environment, and import masks, centroids, polygons, or GeoJSON back into
wsiTools. OME-Zarr metadata reads are lightweight; full chunked pixel access is
best treated as an optional backend rather than a mandatory dependency.

Always check:

```{r, eval=FALSE, purl=FALSE}
wsi_backends()
```

Close slide handles when finished:

```{r, eval=FALSE, purl=FALSE}
wsi_close(slide)
```
