---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## Start With The R Package

Install the lightweight R package first. Optional image and analysis tools are
detected at runtime and are not silently bundled.

```{r, purl = FALSE}
install.packages("remotes", repos = "https://cloud.r-project.org")

remotes::install_github(
  "tkcaccia/wsiTools",
  upgrade = "never",
  build_vignettes = FALSE
)

library(wsiTools)
wsi_backends()
wsi_diagnose(live_test = FALSE)
```

## Optional Tools

Use `wsi_setup()` to print a CRAN-safe setup report. It does not install
anything unless `install = TRUE` is supplied.

```{r, purl = FALSE}
wsi_setup()
wsi_install_backends(install = FALSE)
```

Common optional tools are:

- libvips for conversion, pyramids, thumbnails, and fast tiles.
- OpenSlide for many digital pathology WSI formats.
- native CZI or Bio-Formats for microscopy CZI workflows.
- `httpuv` for live browser-to-R synchronization.
- StarDist or Mesmer only when selected-ROI cell segmentation is explicitly
  configured.

## More Detail

The full installation notes are maintained in the repository:

- [Installation guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/installation.md)
- [Backend setup guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/backends.md)
- [Troubleshooting](https://github.com/tkcaccia/wsiTools/blob/main/docs/troubleshooting.md)
