---
title: "Live viewer security and data handling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Live viewer security and data handling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This page describes the trust boundary of a live wsiTools viewer. It is part
of the reproducibility record for analyses that use `wsi_viewer_live()` or the
Tauri desktop launcher.

## What runs where

The browser renders the interface and requests image tiles. R owns the slide
handle, the optional tile service, project state, annotations, and analysis
objects. The browser communicates with R through a local `httpuv` service using
WebSocket transport when available and HTTP polling as a fallback.

The browser receives data explicitly returned by the viewer session. It does
not receive an R console, an `eval()` endpoint, or a general-purpose command
interpreter. Incoming messages are parsed as structured JSON and are checked
against an event and payload allowlist before they are applied to the R-side
state.

```{r, eval=FALSE, purl=FALSE}
library(wsiTools)

slide <- wsi_open("/path/to/image.svs")
viewer <- wsi_viewer_live(
  slide,
  host = "127.0.0.1",
  transport = "auto",
  tiled = TRUE
)
viewer$open()
```

The default host is loopback (`127.0.0.1`). This keeps the bridge local to the
computer running R. The printed URL contains the local port and a session
identifier; use that URL rather than opening the generated HTML with a
`file://` URL.

## What the browser is allowed to request

The live bridge supports a finite set of viewer events, including viewport
changes, ROI edits, layer and channel changes, project actions, segmentation
status, and spatial selections. State payloads also have a finite set of
accepted fields. Unknown event names, unknown top-level state fields, invalid
object types, and malformed numeric fields are rejected with an actionable R
error.

This validation is an interoperability contract, not an authorization system.
It prevents accidental or unsupported state mutation, but it should not be
treated as a way to expose the viewer to an untrusted network.

## Local-only usage is the supported security model

Do not bind a live viewer to `0.0.0.0`, a public interface, or a forwarded port
unless you have placed it behind an independently managed authentication and
TLS boundary. The live bridge is designed for a trusted local desktop or
remote session in which the browser and R process are on the same protected
machine.

In particular:

- do not paste a live viewer URL into a public issue, chat, notebook, or web
  page;
- do not expose a live viewer while the source slide contains identifiable or
  confidential data;
- stop the viewer when the analysis is finished;
- keep project, log, tile-cache, and exported-file directories access
  controlled by the operating system;
- treat browser logs and saved project files as potentially sensitive because
  they can contain image paths, object identifiers, and annotation metadata.

```{r, eval=FALSE, purl=FALSE}
# Stop the local bridge when finished.
viewer$stop()

# Inspect the bridge and backend status without opening a slide.
wsi_diagnose(live_test = FALSE)
```

## File and memory handling

The core viewer is tile-based. It requests bounded regions through the selected
backend instead of loading a whole WSI into the browser or R memory as one
large raster. Dynamic tiles and temporary caches are local runtime artifacts;
they should be placed on a protected disk and removed when no longer needed.

Project files generally store paths and analysis state rather than a copy of
the original whole-slide pixels. A project is therefore reproducible only when
the source files, checksums or stable accessions, package version, backend
versions, and transforms are retained together.

## Reporting checklist

For a paper, release, or incident report, record:

1. the exact wsiTools release or commit;
2. R, operating-system, browser, and optional backend versions;
3. the host and transport mode (`127.0.0.1`, WebSocket or polling);
4. whether dynamic tiles, static tiles, or a desktop launcher was used;
5. project and tile-cache locations without publishing confidential paths;
6. whether any non-default host, port forwarding, or reverse proxy was used;
7. the `wsi_backends()` and `wsi_diagnose(live_test = FALSE)` output;
8. the saved project, event history, and relevant logs after redacting private
   identifiers.

The live bridge is an R-to-browser integration layer. It is not a network
service, access-control layer, or replacement for an authenticated pathology
data platform.

