---
title: "Segmentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Segmentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## Optional Engines

wsiTools keeps segmentation tools optional. The core package can import cell
masks, centroids, polygons, or GeoJSON without installing StarDist or Mesmer.

```{r, purl = FALSE}
wsi_cell_segmentation_engines()
wsi_has_stardist()
wsi_has_mesmer()
```

## Selected-ROI Workflow

When an external command is configured, use selected ROI crops rather than
processing the whole slide.

```{r, purl = FALSE}
viewer <- wsi_viewer_live(slide, tiled = TRUE)
viewer$open()

roi <- viewer$get_selected_roi()
crop <- export_roi_crop(slide, roi, file = "roi_crop.tif")
```

## Import Existing Results

```{r, purl = FALSE}
cells <- import_segmentation("cells.geojson")
viewer$add_segmentation(cells)
viewer$get_segmentation()
```

More detail:

- [Cell segmentation guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/cell-segmentation.md)
