---
title: "Spatial omics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial omics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## Supported Object Types

wsiTools provides a shared viewer architecture for spatial objects, including:

- Seurat
- Giotto
- SpatialExperiment
- CellPhenotyper project outputs

The goal is to keep high-resolution image pixels on disk while synchronizing
spots, annotations, selections, and derived tables with R.

## Seurat Example

```{r, purl = FALSE}
linked <- wsi_link_seurat_image(
  object = brain,
  image = "/path/to/high_resolution_image.tif",
  image_id = "anterior1",
  coordinate_transform = list(flip = "vertical", rotate = 90)
)

viewer <- wsi_viewer_seurat(
  brain,
  "/path/to/high_resolution_image.tif",
  linked = linked
)
viewer$open()

viewer$colour_spots_by_gene("Mbp")
viewer$get_selected_spots()
viewer$get_spot_annotation_table()
```

## Giotto And SpatialExperiment

```{r, purl = FALSE}
giotto_linked <- wsi_link_giotto_image(giotto_object, image = "slide.tif")
spe_linked <- wsi_link_spatialexperiment_image(spe_object, image = "slide.tif")
```

More detail:

- [Spatial omics guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/spatial-omics.md)
