---
title: "Tile extraction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tile extraction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## Coordinate Manifests First

Tile grids are coordinate manifests. They do not read image pixels.

```{r, purl = FALSE}
grid <- wsi_tile_grid(
  slide,
  tile_size = 512,
  overlap = 0,
  level = 0
)

head(grid)
```

## Extract Tiles

Tiles can be extracted from the whole image, an ROI, or a list of coordinates.
The manifest records provenance columns such as tile coordinates, size, level,
class labels, tissue fraction, and output paths.

```{r, purl = FALSE}
manifest <- extract_tiles(
  slide,
  roi = rois,
  output_dir = "tiles",
  tile_size = 512,
  stride = 512,
  skip_background = TRUE,
  split = c(train = 0.7, valid = 0.3),
  seed = 1
)
```

## Viewer Preview

In live viewers, tile grids can be previewed before extraction so users can
confirm the exported regions visually.

```{r, purl = FALSE}
viewer$preview_tiles(grid)
viewer$extract_tile_preview(output_dir = "tiles")
```

More detail:

- [Examples gallery](https://github.com/tkcaccia/wsiTools/blob/main/docs/examples.md)
