---
title: "Troubleshooting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Troubleshooting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

Start here when wsiTools does not install, an image does not open, the viewer is
white, live synchronization fails, or annotations do not align.

## First Diagnostics

Run these commands and keep the output:

```{r, purl = FALSE}
library(wsiTools)

packageVersion("wsiTools")
wsi_backends()
wsi_setup_report()
wsi_diagnose()
```

For a live viewer session:

```{r, purl = FALSE}
viewer$capabilities()
viewer$get_state()
viewer$get_performance()
```

## Common Problems

| Problem | First check |
| --- | --- |
| `there is no package called 'wsiTools'` | reinstall from GitHub and check `.libPaths()` |
| Windows `00LOCK-wsiTools` | remove the stale `00LOCK-wsiTools` directory |
| Windows `Makeconf:296` | check Rtools and C++ compilation output |
| CZI opens but no image appears | check native CZI or Bio-Formats availability |
| Image is slow or tiles fail | check libvips/OpenSlide/native CZI and tile cache |
| Viewer opens as `file://` | use `wsi_viewer_live()` and the `http://127.0.0.1` URL |
| R sync says unsupported field/event | update wsiTools and regenerate the viewer |
| Scale bar unavailable | check slide metadata with `wsi_mpp()` |
| Seurat/Giotto/SpatialExperiment coordinates are misaligned | check sample mapping, image IDs, flips, rotations, and saved registration |

## Save A Report

For remote debugging, include:

```{r, purl = FALSE}
sink("wsiTools-diagnose.txt")
sessionInfo()
wsi_backends()
wsi_diagnose()
sink()
```

In the viewer, use the Logs or History panel to copy the browser-side report.

The long troubleshooting guide is maintained at:

- [Troubleshooting guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/troubleshooting.md)

