---
title: "Tutorials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This page is the practical entry point for wsiTools. It is organized by the
question a user is likely to have, rather than by function name.

## First 5 Minutes

Install the package, check the available image backends, and open one image.

```{r, purl = FALSE}
install.packages("remotes", repos = "https://cloud.r-project.org")
remotes::install_github("tkcaccia/wsiTools", upgrade = "never")

library(wsiTools)
wsi_backends()

viewer <- wsi_open_viewer("/path/to/image.svs")
```

Use the desktop app if you prefer a file picker instead of R commands. The app
still starts R in the background and opens the same live viewer.

## Viewer Tutorials

| Task | Start here |
| --- | --- |
| Open one SVS, TIFF, BTF, OME-TIFF, or CZI image | [Open one image](https://github.com/tkcaccia/wsiTools/blob/main/docs/open-one-image.md) |
| Understand static versus live viewing | [Viewer](viewer.html) and [live viewer guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/live-viewer.md) |
| Diagnose missing tiles, black viewer, or slow CZI files | [Troubleshooting](troubleshooting.html) |
| Use the optional desktop app | [Desktop app](desktop-app.html) |
| Compare multiple images side by side | [Viewer](viewer.html#full-resolution-viewing) |

## Annotation Tutorials

| Task | Start here |
| --- | --- |
| Draw tissue ROIs | [Annotations](annotations.html) |
| Import or export QuPath GeoJSON | [Annotations](annotations.html#qupath-compatible-geojson) |
| Convert masks to ROIs or ROIs to masks | [Annotations](annotations.html#masks-and-geometry) |
| Associate spots or cells to annotations | [Spatial omics](spatial-omics.html) |
| Keep annotation coordinates aligned after zooming | [Concepts](concepts.html#coordinate-systems) |

## Spatial Omics Tutorials

| Task | Start here |
| --- | --- |
| Open a Seurat object with a high-resolution image | [Spatial omics](spatial-omics.html#seurat) |
| Open a Giotto project | [Spatial omics](spatial-omics.html#giotto-and-spatialexperiment) |
| Open a SpatialExperiment project | [Spatial omics](spatial-omics.html#giotto-and-spatialexperiment) |
| Work with multiple tissue sections | [Spatial omics](spatial-omics.html) |
| Manually register spatial coordinates | [Spatial omics](spatial-omics.html) |
| Colour cells or spots by gene | [Spatial omics](spatial-omics.html) |

## CellPhenotyper Tutorials

| Task | Start here |
| --- | --- |
| Open a CellPhenotyper output folder | [CellPhenotyper](cellphenotyper.html) |
| Overlay H&E and GigaTIME or mIHC channels | [CellPhenotyper](cellphenotyper.html#typical-layers) |
| Load cell masks, cell centroids, or KODAMA regions | [CellPhenotyper](cellphenotyper.html) |
| Run proximity or trajectory analysis | [Spatial omics](spatial-omics.html) |

## Copy-Paste Examples

Installed example scripts are available with `system.file()`:

```{r, purl = FALSE}
system.file("examples/open_single_svs_live.R", package = "wsiTools")
system.file("examples/open_czi_project_live.R", package = "wsiTools")
system.file("examples/open_spatialexperiment_four_slide_live.R", package = "wsiTools")
system.file("examples/open_cellphenotyper_project_live.R", package = "wsiTools")
system.file("examples/open_he_mihc_overlay_live.R", package = "wsiTools")
```

For a complete list, see the
[examples gallery](https://github.com/tkcaccia/wsiTools/blob/main/docs/examples.md).

