---
title: "Validation and reproducibility"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation and reproducibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

This page defines what is tested automatically and what should be measured when
reporting a wsiTools result. It is deliberately explicit: a passing R package
check validates software behavior, but it does not by itself establish image
registration accuracy, viewer speed, or biological validity.

## Current validation layers

The repository contains automated tests for several core contracts:

| Layer | Examples | Interpretation |
| --- | --- | --- |
| Package installation | CRAN-safe imports, native registration, optional dependencies | The lightweight package can be built without optional image/model tools |
| Image access | metadata, region reads, thumbnails, tiles, format fallbacks | A backend can satisfy a defined image-access contract for the tested fixture |
| Annotation data | GeoJSON, masks, polygons, multipolygons, class behavior | Data transformations preserve the tested coordinates and metadata |
| Viewer protocol | state payloads, event allowlists, project state, live-session helpers | Browser/R messages follow the tested schema |
| Spatial objects | Seurat, Giotto, SpatialExperiment, cell/spot association | The tested object representation can be read and mapped to image coordinates |

The current evidence status is intentionally separated from implementation
status:

| Claim | Current evidence | What a publication should add |
| --- | --- | --- |
| Lightweight package installation | Clean source build and automated tests | Repeated checks on supported R/platform versions |
| Core annotation and GeoJSON behavior | Deterministic unit-test fixtures | Large, representative annotations and round-trip error measurements |
| Live viewer protocol | Event/state tests and manual smoke workflows | Reconnection, failure, and multi-platform latency measurements |
| Large-image performance | Instrumentation and performance report API | Cold/warm-cache benchmarks on real slides against comparator tools |
| Spatial-coordinate integration | Object-specific tests and explicit transforms | Landmark error and section-level registration validation |
| Biological analysis | Workflow interfaces and import/export paths | A biological case study with independent validation |

This matrix is part of the documentation rather than a claim that all rows have
already been validated experimentally.

Run the package tests locally with:

```{r, purl = FALSE}
testthat::test_local()
```

For a release-level check, use a clean source tree without private benchmark
data and run:

```{bash, eval=FALSE}
R CMD build .
R CMD check --as-cran wsiTools_<version>.tar.gz
```

Optional backends are capability-tested and skipped when unavailable. A green
check therefore means that the tested core is healthy; it does not mean that
every CZI, SVS, BTF, Bio-Formats, StarDist, or Mesmer installation is supported.

## Performance benchmark protocol

For a reproducible performance report, record the following for every image and
backend combination:

1. file format, scanner/vendor, dimensions, pyramid levels, and file size;
2. operating system, CPU, RAM, R version, wsiTools version, and backend versions;
3. time to metadata, time to first preview, time to first requested tile, and
   time to a full-viewport redraw;
4. peak resident memory and CPU usage during startup, zooming, and panning;
5. tile cache mode, cache size, tile dimensions, and cache hit rate;
6. number of tile failures and retry behavior;
7. whether the viewer was static, polling-live, or WebSocket-live.

Use the viewer performance report as part of the record:

```{r, purl = FALSE}
viewer$get_performance()
wsi_diagnose(live_test = FALSE)
```

Do not compare timings from a cold cache with timings from a warm cache. Report
both when cache behavior is relevant. Do not describe a workflow as
“full-resolution” without stating the physical pixel size or pyramid level that
was displayed.

## Registration validation

Spatial overlays should be validated independently of visual impression. For
each image/object pair, retain:

- the original object coordinates;
- the image identifier and section identifier;
- the flip, rotation, scale, translation, or registration transform;
- the number of coordinates retained and excluded;
- landmark coordinates before and after transformation;
- residual error in pixels and, when available, micrometres.

For a practical landmark check, select corresponding tissue landmarks in the
image and spatial object, then report the median and 95th-percentile Euclidean
error. If a transform is estimated from landmarks, keep the landmarks and the
transform in project provenance rather than only saving the transformed object.

```{r, purl = FALSE}
# Inspect the coordinates and image metadata before registration.
coords <- wsi_seurat_coordinates(object, image = "slice1")
info <- wsi_info(slide)
info
head(coords)
```

Visual agreement alone is not sufficient evidence of registration accuracy,
especially for objects containing multiple tissue sections or cell-level and
spot-level coordinates at the same time.

## Annotation validation

For annotation workflows, validate both the representation and the geometry:

- GeoJSON import followed by export should preserve class, label, object ID, and
  slide-coordinate geometry;
- mask conversion should record pixel size, origin, label mapping, and whether
  holes or small components were removed;
- class-exclusive editing should report how intersections are resolved;
- spot/cell association should record the point-in-polygon rule and boundary
  convention;
- reopening the saved project should reproduce the same annotation coordinates.

The package tests cover small deterministic fixtures. For a scientific report,
also include a representative large annotation set and report vertex count,
rendering time, memory use, and the geometry difference introduced by any
simplification or rasterization.

## Reproducible project records

Save the project together with a diagnostic report and a session record:

```{r, purl = FALSE}
project <- wsi_project(slide)
wsi_save_project(project, "case_01.wsiproject")

writeLines(capture.output(sessionInfo()), "case_01-sessionInfo.txt")
writeLines(capture.output(wsi_backends()), "case_01-backends.txt")
writeLines(capture.output(wsi_diagnose(live_test = FALSE)), "case_01-diagnose.txt")
```

Project files store references and analysis state; they do not generally copy
the original whole-slide pixels. A reproducible record must therefore retain
the source-image paths or stable data-access instructions, file checksums when
possible, the object files, backend versions, and the exact wsiTools release.

## Scope and current limitations

wsiTools is an evolving research tool. The following should be treated as
backend- or dataset-dependent until a corresponding benchmark is published:

- universal support for every CZI, BTF, OME-TIFF, SVS, or DICOM WSI variant;
- performance equivalence between native CZI, libvips, OpenSlide, and
  Bio-Formats;
- automatic registration of arbitrary spatial objects to arbitrary microscopy
  images;
- biological correctness of externally generated segmentation or stain
  deconvolution;
- synchronization after browser refresh, network interruption, or stale R
  sessions;
- exact equivalence between vector geometry and raster mask rendering.

When reporting a result, state the tested file format, backend, operating system,
package version, and any manual transformation or correction applied.
