---
title: "Viewer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Viewer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, purl = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

## Static Viewer

Use a static viewer when you want a portable HTML page and do not need browser
edits to return automatically to R.

```{r, purl = FALSE}
library(wsiTools)

slide <- wsi_open("/path/to/image.svs")
wsi_viewer(slide)
```

Static viewers may use precomputed Deep Zoom tiles for smooth full-resolution
zooming.

## Live Viewer

Use a live viewer when annotations, selections, measurements, trajectories,
channels, or project state should come back to the active R session.

```{r, purl = FALSE}
viewer <- wsi_viewer_live(
  slide,
  tiled = TRUE
)

viewer$open()
viewer$get_rois()
viewer$get_selected_roi()
viewer$get_measurements()
viewer$get_channel_settings()
```

The live viewer opens an `http://127.0.0.1:<port>/...` URL through `httpuv`.
If the page is opened as `file://`, automatic synchronization with R is not
active.

## Full-Resolution Viewing

The browser needs image tiles, not raw SVS, CZI, BTF, or OME-TIFF file paths.
wsiTools can use:

- prebuilt Deep Zoom tiles for the smoothest static viewing;
- dynamic live tile endpoints for on-demand region reads;
- OpenSeadragon overlays for annotations, spots, masks, and channels.

More detail:

- [Live viewer guide](https://github.com/tkcaccia/wsiTools/blob/main/docs/live-viewer.md)
- [Architecture diagram](https://github.com/tkcaccia/wsiTools/blob/main/docs/architecture.md)
- [Visual gallery](https://github.com/tkcaccia/wsiTools/blob/main/docs/screenshots.md)
